Moment Distances from Robust Subspace for Botnet Detection

### 2019-05-17 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on results

We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}. Finally, We evaluate the accuracy of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.

@@

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized. Then, Section \ref{sec:results} shows the results of the proposed approach and compares to selected anomaly detection algorithms.

@@

Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training the model, while the latter does not require labeled data or training. However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or unkown events.

Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches. The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomalous detection, respectively.

@@

This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches. As clustering-based method, we compare to K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}. It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, and that algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}. Therefore, we compare our proposal to GMM. We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.

In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach for all botnet malwares of the CTU-13. Therefore, the data set and scenarios for our evaluation can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, ... , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:tab01}.

Given a contaminated data set $\boldsymbol{X}$, split it into $\boldsymbol{X}^n$ containing 60\% of the the normal data, and into cross validation $\boldsymbol{X}^v$, composed by 20\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}^t$ containing 20\% of the normal and 50\% of the anomalous data.

Evaluate if the code is correct and implement it as funcions in my code

try other distributions between train cv and test

Explain tha numbers of anomalies and normal for CV and test

Select the range of contamination that shall be evaluated and perform cross validation for each contamination.

For each contamination $c$:

Estimate $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$ from normal data.

Predict anomalies for cross validation data using the $c$, $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.

Select the best F-measure by cross validation and get the best $c$.

Test the F-measure of anomaly detection for new observations through $d(\boldsymbol{x},\bar{\boldsymbol{x}}_0, \boldsymbol{\hat{S}}_0)$ and $c$.

\cite{garcia2014empirical} also uses a time window separation and an aggregation method.

\textbf{Experiment Plan:}

Evaluation Measure:

In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply an accuracy evaluation, we would have high accuracy but poor true-positive detection.

Due to the importance given by F-measure to true-positive detection \cite{powers2011evaluation,moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.

\textbf{Experiment Plan:}

	The F-measure is the harmonic mean of precision and recall and is defined by 
	\begin{equation}\label{eq:eq15}
		F-measure = 2 * \frac{PR * RC}{PR + RC},				
	\end{equation}
	where $PR$ is the precision, defined by 
	\begin{equation}\label{eq:eq16}
		PR = \frac{True Positive}{True Positive + False Positive}
	\end{equation}
	and $RC$ is the recall, defined by 
	\begin{equation}\label{eq:eq17}
		RC = \frac{True Positive}{True Positive + False Negative}.
	\end{equation}

\textbf{Data splitting for training, cross validation and testing}

@@

\begin{table}[h!]
  \centering
  \scriptsize
  \caption{Results grouped by algorithm and scenario}
  \label{tab:tab02}
  \begin{tabular}{ c|c c|c c|c c|c c|c c }
	\toprule
	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{S-MCD}} &\multicolumn{2}{c}{\textbf{K-MCD}} &\multicolumn{2}{c}{\textbf{MCD}} &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}}\\ 
			\hhline{~----------}
			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
	\midrule
		10 &\color{red} 0.71 & 0.03 & 0.66 & 0.02 & 0.34 & 0.52 & 0.34 & 0.46 & 0.20 & 0.25 \\ \hline
		11 & 0.03 & 0.54 & 0.00 & 0.54 & 0.71 &\color{red} 0.73 & 0.35 & 0.48 & 0.13 & 0.15 \\ \hline
		12 & 0.19 & 0.02 & 0.20 & 0.05 & 0.06 &\color{red} 0.31 & 0.01 & 0.01 & 0.09 & 0.11 \\ \hline
		15 & 0.00 & 0.00 & 0.00 & 0.00 & 0.24 &\color{red} 0.32 & 0.14 & 0.22 & 0.09 & 0.11 \\ \hline
		15-2 & 0.04 & 0.85 & 0.67 &\color{red} 0.86 & 0.36 & 0.28 & 0.32 & 0.14 & 0.18 & 0.26 \\ \hline
		15-3 & 0.00 & 0.01 & 0.00 & 0.01 &\color{red} 0.55 & 0.48 & 0.00 & 0.01 & 0.22 & 0.29 \\ \hline
		16 & 0.45 & 0.36 & 0.42 & 0.33 & 0.27 & 0.41 &\color{red} 0.51 & 0.49 & 0.07 & 0.16 \\ \hline
		16-2 & 0.00 & 0.16 & 0.00 &\color{red} 0.20 & 0.04 & 0.04 & 0.02 & 0.08 & 0.04 & 0.03 \\ \hline
		16-3 & 0.00 & 0.05 & 0.00 & 0.07 &\color{red} 0.15 & 0.10 & 0.01 & 0.01 & 0.03 & 0.04 \\ \hline
		17 & 0.03 & 0.03 & 0.07 & 0.03 &\color{red} 0.59 & 0.53 & 0.16 & 0.47 & 0.35 & 0.39 \\ \hline
		18 & 0.63 & 0.10 & 0.63 & 0.08 & 0.70 &\color{red} 0.72 & 0.21 & 0.17 & 0.13 & 0.20 \\ \hline
		18-2 & 0.83 & 0.69 & 0.85 & 0.43 & 0.84 &\color{red} 0.91 & 0.50 & 0.55 & 0.62 & 0.70 \\ \hline
		19 &\color{red} 0.71 & 0.00 & 0.68 & 0.00 & 0.11 & 0.42 & 0.17 & 0.20 & 0.20 & 0.17 \\ \hline
		\rowcolor{Gray} \textbf{avg} & \textbf{0.28} & \textbf{0.22} & \textbf{0.32} & \textbf{0.20} & \textbf{0.38} &\color{red} \textbf{0.44} & \textbf{0.21} & \textbf{0.25} & \textbf{0.18} & \textbf{0.22} \\ 
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h!]
  \centering
  \scriptsize
  \caption{Results grouped by algorithm and scenario}
  \label{tab:tab03}
  \begin{tabular}{ c|c c|c c|c c|c c }
	\toprule
	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA2}} &\multicolumn{2}{c}{\textbf{MCD}}\\ 
			\hhline{~--------}
			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
	\midrule
		10 & 0.93 & 0.93 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.34 & 0.52 \\ \hline
		11 & 0.96 & 0.94 &\color{red} 0.99 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.71 & 0.73 \\ \hline
		12 & 0.32 & 0.19 & 0.00 & 0.00 &\color{red} 0.81 & 0.01 & 0.06 & 0.31 \\ \hline
		15 & 0.75 & 0.73 & 0.86 &\color{red} 0.96 &\color{red} 1.00 &\color{red} 0.96 & 0.24 & 0.32 \\ \hline
		15-2 & 0.40 & 0.57 &\color{red} 0.95 &\color{red} 0.94 & 0.25 & 0.93 & 0.36 & 0.28 \\ \hline
		15-3 & 0.68 & 0.68 &\color{red} 0.96 & 0.70 &\color{red} 0.96 &\color{red} 0.96 & 0.55 & 0.48 \\ \hline
		16 &\color{red} 0.47 &\color{red} 0.47 & 0.01 & 0.01 & 0.01 & 0.00 & 0.27 & 0.41 \\ \hline
		16-2 & 0.07 & 0.10 &\color{red} 0.80 & 0.68 & 0.00 & 0.00 & 0.04 & 0.04 \\ \hline
		16-3 & 0.29 & 0.29 & 0.01 & 0.46 &\color{red} 0.79 & 0.76 & 0.15 & 0.10 \\ \hline
		17 &\color{red} 0.82 & 0.79 &\color{red} 0.81 & 0.78 &\color{red} 0.82 & 0.79 & 0.59 & 0.53 \\ \hline
		18 & 0.87 & 0.89 & 0.97 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.98 & 0.70 & 0.72 \\ \hline
		18-2 &\color{red} 0.94 &\color{red} 0.94 &\color{red} 0.91 & 0.86 & 0.09 & 0.10 & 0.84 & 0.91 \\ \hline
		19 & 0.27 & 0.31 &\color{red} 0.95 &\color{red} 0.95 & 0.02 & 0.02 & 0.11 & 0.42 \\ \hline
		\rowcolor{Gray} \textbf{Avg} & \textbf{0.60} & \textbf{0.60} &\color{red} \textbf{0.71} &\color{red} \textbf{0.71} & \textbf{0.59} & \textbf{0.58} & \textbf{0.38} & \textbf{0.44} \\
    \bottomrule
  \end{tabular}
\end{table}

Results grouped by algorithm for full CTU-13.

\subsection{Discussion}
\label{sec:discussion}

@@

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

@@

% Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,...,\boldsymbol{x}_n\}$ of $p$-variate observations. Let $\boldsymbol{h} \subset \{1,...,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.

@@

% for $i = 1,...,n$.

% Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,...,n$.

% Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},...,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq ... \leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.

@@

% \textbf{The C-step can be algorithmically described as follows:}
% \begin{algorithm}
% 	\label{alg:alg01}
% 	\scriptsize
% 	\SetAlgoLined
% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
% 	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
% 		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
% 			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
% 		}
% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
% 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
% 		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
% 	}
% 	\caption{C-Step}
% \end{algorithm}

@@

% \textbf{Initial subset can be algorithmically described as:}
% \begin{algorithm}
% 	\label{alg:alg02}
% 	\scriptsize
% 	\SetAlgoLined
% 	\KwResult{$\boldsymbol{h}_1$}
% 	Given a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $p$-variate observations\;
% 	Draw a random $(p + 1)$-subset $\boldsymbol{L}$\;
% 	Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
% 	\If{$det(\boldsymbol{\hat{S}}_1) \leq 0$}{
% 		Add another random observation into $\boldsymbol{L}$\;
% 		Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
% 	}
% 	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, ..., n$\;
% 	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_0(\pi(n))$\;
% 	Put $\boldsymbol{h}_1 = \{\pi(1), ..., \pi(h)\}$\;
% 	\caption{Constructing the initial subset}
% \end{algorithm}

>>>>

We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}. Finally, We evaluate the F-measure of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.

@@

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized. Section \ref{sec:results} shows the results of the proposed approach, compares to selected anomaly detection algorithms, and presents a discussion about the results.

@@

Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training anomaly detection models, while the latter does not require labeled data or training. However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unkown events.

Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches. The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomaly detection, respectively.

@@

This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches. As clustering-based method, we compare to K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}. It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, and that algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}. Therefore, we compare our proposal to GMM. We also evaluate the results of ROBPCA for anomaly detection on CTU-13, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.

In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach to all botnet malwares of the CTU-13.

Therefore, the full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \\\ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:tab01}. In our experiment setup each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into the cross validation $\boldsymbol{X}_i^v$, composed by 25\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}_i^t$ containing 25\% of the normal and 50\% of the anomalous data.

In anomalous detection problems, where anomalies are rare events, if one classify all observations as normal and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection. Due to the importance given by the F-measure to true-positive detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic} and it is the metric used for validation of this experiment.

The F-measure is the harmonic mean of precision and recall and is defined by:
\begin{equation}\label{eq:eq15}
	F-measure = 2 * \frac{pr * rc}{pr + rc},
\end{equation}
where $pr$ is the precision, defined by 
\begin{equation}\label{eq:eq16}
	pr = \frac{True Positive}{True Positive + False Positive}
\end{equation}
and $rc$ is the recall, defined by 
\begin{equation}\label{eq:eq17}
	rc = \frac{True Positive}{True Positive + False Negative}.
\end{equation}

It is from $\boldsymbol{X}_i^n$ that we perform the training steps, which are the steps between 1 and 10 described in Algorithm \ref{alg:alg01}, resulting into $\boldsymbol{\bar{x}}$, $\boldsymbol{\bar{\Sigma}}$, $\boldsymbol{\bar{s}}$ and $\boldsymbol{\bar{k}}$. The next step of the experiment is the cross validation, which is an empirical testing to identify the contamination $c$ that obtains the best F-measure on anomaly detection, aiming the maximum precision and recall. During the cross validation for $\boldsymbol{X}_i^v$ we compute the steps 11, 12 and 13, and iterate over the steps between 14 and 16 for selected values of $c$, and validate the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}^v$ to the ground truth results for $\boldsymbol{X}_i^v$, in order to identify the $c$ that best classify the anomalies of $\boldsymbol{X}_i^v$.

Finally, we test the proposed approach to anomaly detection on $\boldsymbol{X}_i^t$ by means of the steps between 11 and 16 of Algorithm \ref{alg:alg01}, and compare the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$ to the ground truth results for $\boldsymbol{X}_i^t$.

@@

The Table \ref{tab:tab02} presents the results of the proposed approach and selected methods applied to botnet detection for each scenario of the CTU-13 data set. The Table \ref{tab:tab02} shows the F-measure of GMM, K-means, ROBPCA, md-RPCA, kd-RPCA and sd-RPCA by scenario and by an average of all scenarios.

\begin{table}[h!]
  \centering
  \scriptsize
  \caption{Results grouped by algorithm and scenario}
  \label{tab:tab02}
  \begin{tabular}{ c|c c|c c|c c|c c|c c|c c }
	\toprule
	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}} &\multicolumn{2}{c}{\textbf{ROBPCA}} &\multicolumn{2}{c}{\textbf{md-RPCA}} &\multicolumn{2}{c}{\textbf{kd-RPCA}} &\multicolumn{2}{c}{\textbf{sd-RPCA}}\\ 
			\hhline{~------------}
			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
	\midrule
		10 & 0.34 & 0.45 & 0.20 & 0.25 & 0.29 & 0.34 & 0.97 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
		11 & 0.34 & 0.47 & 0.12 & 0.14 & 0.27 & 0.37 & 0.98 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
		12 & 0.00 & 0.00 & 0.09 & 0.11 & 0.08 & 0.11 & 0.86 & 0.80 & 0.00 & 0.00 & 0.94 & 0.90 \\ \hline
		15 & 0.14 & 0.21 & 0.08 & 0.11 & 0.05 & 0.07 & 0.85 & 0.79 & 1,00 & 0.96 & 0.99 & 0.96 \\ \hline
		15-2 & 0.32 & 0.14 & 0.18 & 0.25 & 0.21 & 0.30 & 0.85 & 0.82 & 0.98 & 0.99 & 0.98 & 0.99 \\ \hline
		15-3 & 0.00 & 0.01 & 0.22 & 0.28 & 0.28 & 0.38 & 0.95 & 0.96 & 0.00 & 0.00 & 0.99 & 0.98 \\ \hline
		16 & 0.50 & 0.48 & 0.06 & 0.16 & 0.06 & 0.10 & 0.85 & 0.83 & 0.00 & 0.00 & 0.96 & 0.00 \\ \hline
		16-2 & 0.02 & 0.08 & 0.03 & 0.02 & 0.00 & 0.00 & 0.22 & 0.26 & 0.80 & 1,00 & 0.80 & 1,00 \\ \hline
		16-3 & 0.00 & 0.01 & 0.02 & 0.04 & 0.01 & 0.03 & 0.60 & 0.66 & 0.00 & 0.00 & 0.84 & 0.98 \\ \hline
		17 & 0.16 & 0.46 & 0.35 & 0.38 & 0.72 & 0.75 & 0.81 & 0.78 & 0.02 & 0.78 & 0.81 & 0.78 \\ \hline
		18 & 0.21 & 0.17 & 0.13 & 0.19 & 0.13 & 0.15 & 0.91 & 0.88 & 0.97 & 0.98 & 0.97 & 0.98 \\ \hline
		18-2 & 0.50 & 0.54 & 0.61 & 0.70 & 0.27 & 0.24 & 0.80 & 0.85 & 0.99 & 0.98 & 0.99 & 0.98 \\ \hline
		19 & 0.16 & 0.20 & 0.19 & 0.17 & 0.10 & 0.13 & 0.91 & 0.91 & 0.96 & 0.99 & 0.96 & 0.99 \\ \hline
		\rowcolor{Gray} \textbf{avg} &\textbf{0.21} &\textbf{0.25} &\textbf{0.18} & \textbf{0.21} & \textbf{0.19} &\textbf{0.23} &\textbf{0.81} &\textbf{0.80} &\textbf{0.59} &\textbf{0.67} &\textbf{0.94} &\textbf{0.89} \\
    \bottomrule
  \end{tabular}
\end{table}

It is possible to observe that the results of md-RPCA, kd-RPCA and sd-RPCA overcome the remain methods in a comparison by scenario and by average F-measure.

@@

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting F-measure of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

@@

% Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,\ldots,\boldsymbol{x}_n\}$ of $p$-variate observations. Let $\boldsymbol{h} \subset \{1,\ldots,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.

@@

% for $i = 1,\ldots,n$.

% Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,\ldots,n$.

% Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},\ldots,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq \ldots \leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.

@@

% \textbf{The C-step can be algorithmically described as follows:}
% \begin{algorithm}
% 	\label{alg:alg01}
% 	\scriptsize
% 	\SetAlgoLined
% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
% 	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
% 		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
% 			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
% 		}
% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,\ldots, n$\;
% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq \ldots \leq \boldsymbol{d}_1(\pi(n))$\;
% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), \ldots, \pi(h)\}$\;
% 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
% 		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
% 	}
% 	\caption{C-Step}
% \end{algorithm}

@@

% \textbf{Initial subset can be algorithmically described as:}
% \begin{algorithm}
% 	\label{alg:alg02}
% 	\scriptsize
% 	\SetAlgoLined
% 	\KwResult{$\boldsymbol{h}_1$}
% 	Given a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $p$-variate observations\;
% 	Draw a random $(p + 1)$-subset $\boldsymbol{L}$\;
% 	Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
% 	\If{$det(\boldsymbol{\hat{S}}_1) \leq 0$}{
% 		Add another random observation into $\boldsymbol{L}$\;
% 		Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
% 	}
% 	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, \ldots, n$\;
% 	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq \ldots \leq \boldsymbol{d}_0(\pi(n))$\;
% 	Put $\boldsymbol{h}_1 = \{\pi(1), \ldots, \pi(h)\}$\;
% 	\caption{Constructing the initial subset}
% \end{algorithm}
